---
title: "Reproducable Research - Peer Assignment 1"
author: "Levente Szabados"
date: '2015-07-18'
output: html_document
---

# Homework Assignment I. for Reproducible Research
## Initial steps

As a basic setup, the data sources are as follows:


```{r setup, echo=TRUE}
library(knitr)
opts_chunk$set(echo = TRUE)
workdir <-"~/course-work/C5/"
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
filename <- "activity.zip"
```

Our first task for the assigment is to download the datafile from the given [URL](`r url`), and decompress it and check the summary.

Our working directory will be: *`r workdir`*

```{r download data, cache=TRUE}
download.file(url, paste(workdir,filename), method='curl')
data <- read.table(unzip(paste(workdir,filename),"activity.csv"), header=TRUE, sep=",")
summary(data)
```

First we should check the head of the data table to see how the data in it looks like.

```{r head}
head(data)
```

We see, that **NA** is present in many cases for the *steps* variable. We will be forced to use na.rm = TRUE later on.

We have to merge the interval data, so as to get to the daily sums of steps.

```{r merge days}
aggregated_steps <-aggregate(. ~ date,  data=data, FUN=sum, na.rm=TRUE)
head(aggregated_steps)
```

## First part of assignment question

**What is the mean and median of our data?**

In order to answer this, we prepare the histogram of total steps per day and calculate mean, median values.

```{r histogram}
hist(aggregated_steps$steps, main="Histogram of steps taken daily", xlab="Steps", ylab="Days")
mea <- mean(aggregated_steps$steps)
med <- median(aggregated_steps$steps)
options(scipen=999) # disable scientific notation by output.
```

We can conclude, that the **mean** number of steps is $$`r mea`$$ and the **median** is $$`r med`$$.

## Second part of assignment question
**What is the daily activity pattern?

To be able to visualize this, a time series has be created and data has to be averaged out for intervals.
First, let us check, how the notation is for intervals?

```{r interval data printout}
print(data$interval[seq(50)])
```

Luckily, the intervals have unique *id*s, so we can simply average over them.
Let's see!

```{r average steps in intervals}

# and the data aggregation
aggregated_intervals <-aggregate(. ~ interval,  data=data, FUN=mean, na.rm=TRUE)

head(aggregated_intervals)
```

**COOL!** 
By now, we have the data.
Let's take a look, we might get a pattern!

```{r histogram of intervals}

plot(aggregated_intervals$steps,  type="l", xlab="5 minute intervals of the day", ylab="Average steps taken")
ma <- which(aggregated_intervals$steps==max(aggregated_intervals$steps),arr.ind=TRUE)
maxtime<-sprintf("%04d",aggregated_intervals$interval[ma])
maxtime<-format(strptime(maxtime, format="%H%M"), format = "%H:%M")
```

### Conclusion
The time with the maximal average steps taken daily is **`r maxtime`**.

## Third part of Assignment question

Let us first calculate the number of missing values in the original data!

```{r missing number}
missing_data<-sum(is.na(data$steps))
```

The number of missing datapoints: $`r missing_data`$

Let us fill the places with the mean for the day!

```{r fill NA}
filled_data <- data
na_mask<-is.na(filled_data$steps)
intervals_vec <-aggregated_intervals$steps
filled_data$steps <- with(filled_data, ifelse(is.na(steps), intervals_vec , steps))
```

Now let us again look at te data from the mean, median and histogram ppoint of view!

```{r merge days again and calculate}
aggregated_steps_filled <-aggregate(. ~ date,  data=filled_data, FUN=sum, na.rm=TRUE)
hist(aggregated_steps_filled$steps, main="Histogram of steps taken daily - NA filled", xlab="Steps", ylab="Days")
mea2 <- mean(aggregated_steps_filled$steps)
med2 <- median(aggregated_steps_filled$steps)
options(scipen=999) # disable scientific notation by output.
```

We can conclude, that the **mean** number of steps is $$`r mea2`$$ and the **median** is $$`r med2`$$.
Which is compared to the **original mean** of  $$`r mea`$$ and **original median** $$`r med`$$ a **small** difference, and only in the *median*.

## Fourth part of assignment question

We would like to separate weekdays and weekends, and then run some plotting.

Let us first add a weekday column to the data!

```{r adding weekday data}
data$weekday <- weekdays(as.Date(data$date))
```

After this, we can do a filtering into two parts, and then start the plotting.

```{r subset based on weekdays and plot}
weekend = list("szombat","vasárnap")
weekend_data <- subset(data, weekday %in% weekend)
weekend_data$weekday <-NULL
weekday_data <- subset(data, !(weekday %in% weekend))
weekday_data$weekday <-NULL
#head(weekend_data)
aggregated_weekday_intervals <-aggregate(. ~ interval,  data=weekday_data, FUN=mean, na.rm=TRUE)
aggregated_weekend_intervals <-aggregate(. ~ interval,  data=weekend_data, FUN=mean, na.rm=TRUE)
#head(aggregated_weekend_intervals)

par(mfrow=c(2,1), pin=c(6,1.7))
plot(aggregated_weekend_intervals$steps,  type="l", xlab ="", ylab="")

plot(aggregated_weekday_intervals$steps,  type="l", xlab="", ylab="")

```


###Conclusion

We can conclude, that the activity patterns on weekdays and weekendy are different.
